Education Serves as a Deterrent to Criminal Behaviour

### 2024-03-29 paper/paper.qmd
Add Figures 1 and 2 and results section

library(tidyverse)
library(palmerpenguins)
library(haven)
library(ggplot2)
library(dplyr)
library(sf)
```

@@

# Loading the data
figure1_data <- read_csv(here::here("data/raw_data/reproduction_data/figure1_data.csv"))

```
# Introduction

@@

## Sources
This paper replicates and explores various aspects of the original study for this analysis. It is reproduced using several datasets from the replication package provided by Bell, Costa, and Machin from their paper “Why Does Education Reduce Crime?” [@education]. The replication package only holds the cleaned data in multiple files, depending on how they appear on the paper. Figures 1, 2, and 3 are replicated from the original paper for this reproduction. Our reproduction aims to address the following two questions: (1) Do CSL reforms deter individuals from criminal activity, and thus decreasing crime rate in youth? (2) Does this reduction reshape criminal behaviour distributions by age?

Bell et al. do not analyse a single dataset. Instead, we have many different sources provided to us. The ones used to reproduce and compile this paper are as follows:

@@

```{r}
#| echo: false
#| label: fig1
#| fig-cap: US male arrest rates between the ages 15-24 from FBI Uniform Crime Reporting data

@@

#Plots age against cumulative arrests
ggplot(figure1_data, aes(x = as.factor(agegroup), y = cumulative_arrest, 
                       fill = factor(crime_type))) +
  geom_bar(stat = "identity", position = "stack", width = 0.8) + #bar graph
  scale_fill_manual(values = c("black", "lightblue", "darkred"), #colours
                    labels = c("Drugs", "Property", "Violent")) + #legend
  labs(x = "Age", y = "Arrest Rate (Per 1000 Population)", fill = NULL, 
       title = "Cumalitive Arrests By Age and Crime Type") + #titles
  scale_x_discrete(labels = c("15", "16", "17", "18", "19", "20", "21", "22", 
                              "23", "24", "25-29", "30-34", "35-39")) + #labels
  scale_y_continuous(breaks = seq(0, 100, by = 20)) + # y-axis breaks
  theme_minimal() +
  theme(legend.position = "bottom", #legend position
        axis.text.x = element_text(angle = 0, hjust = 0.5), #Centers axis
        plot.title = element_text(hjust = 0.5), #Centers title
        panel.grid.major = element_blank(),  #Removes major grid lines
        panel.grid.minor = element_blank())  #Removes minor grid lines

@@

In analyzing @fig1, we can see a jump in male arrests for ages 15 to 18. The arrests peak at 18 and then fall gradually thereafter. These are the ages individuals are typically meant to be in high school. Within this figure, we can see how arrest rates vary across different age groups. An analysis can tell us the consistency across age cohorts. The types of crime driving the trend, is of particular interest to us. We observed there is an increase in drug-related crimes for individuals in this age range and is more pronounced that the other two types of crimes. It prompts further investigation and we provide some insight to this in [@sec-results].

@@

>>>>

library(tidyverse)
library(palmerpenguins)
library(haven)
library(ggplot2)
library(dplyr)
library(sf)
library(gridExtra)
```

@@

# Loading the data
figure1_data <- read_csv(here::here("data/raw_data/reproduction_data/figure1_data.csv"))
maps_d <- read_csv(here::here("data/raw_data/reproduction_data/maps_d.csv")) 
maps_d_clean <- read_csv(here::here("data/analysis_data/maps_d_clean.csv"))
dropout_age <- read_csv(here::here("data/raw_data/reproduction_data/dropout_age.csv"))

# Introduction

@@

## Sources
This paper replicates and explores various aspects of the original study for this analysis. It is reproduced using several data sets from the replication package provided with the paper “Why Does Education Reduce Crime?” [@education]. The replication package only holds the cleaned data in multiple files, depending on how they appear on the paper. Figures 1, 2, and 3 are replicated from the original paper for this reproduction. Our reproduction aims to address the following two questions: (1) Do CSL reforms deter individuals from criminal activity, and thus decreasing crime rate in youth? (2) Does this reduction reshape criminal behaviour distributions by age?

@education do not analyse a single dataset. Instead, we have many different sources provided to us. The ones used to reproduce and compile this paper are as follows:

@@

```{r}
#| echo: false
#| label: fig-arrests
#| warning: false
#| fig-cap: US male arrest rates between the ages 15-24 from FBI Uniform Crime Reporting data

@@

#Plots age against cumulative arrests
ggplot(figure1_data, aes(x = as.factor(agegroup), y = cumulative_arrest, 
                       fill = factor(crime_type))) +
  geom_bar(stat = "identity", position = "stack", width = 0.8) + #bar graph
  scale_fill_manual(values = c("black", "lightblue", "darkred"), #colours
                    labels = c("Drugs", "Property", "Violent")) + #legend
  labs(x = "Age", y = "Arrest Rate (Per 1000 Population)", fill = NULL, 
       title = "Cumalitive Arrests By Age and Crime Type") + #titles
  scale_x_discrete(labels = c("15", "16", "17", "18", "19", "20", "21", "22", 
                              "23", "24", "25-29", "30-34", "35-39")) + #labels
  scale_y_continuous(breaks = seq(0, 100, by = 20)) + # y-axis breaks
  theme_minimal() +
  theme(legend.position = "bottom", #legend position
        axis.text.x = element_text(angle = 0, hjust = 0.5), #Centers axis
        plot.title = element_text(size = 12, hjust = 0.5), #Centers title
        panel.grid.major = element_blank(),  #Removes major grid lines
        panel.grid.minor = element_blank())  #Removes minor grid lines

@@

In analyzing @fig-arrests, we can see a jump in male arrests for ages 15 to 18. The arrests peak at 18 and then fall gradually thereafter. These are the ages individuals are typically meant to be in high school. Within this figure, we can see how arrest rates vary across different age groups. An analysis can tell us the consistency across age cohorts. The types of crime driving the trend, is of particular interest to us. We observed there is an increase in drug-related crimes for individuals in this age range and is more pronounced that the other two types of crimes. It prompts further investigation and we provide some insight to this in [@sec-results].

@@

```{r}
#| echo: false
#| label: fig-dropout
#| warning: false
#| message: false
#| fig-cap: Change in dropout ages designated by state law between 1980 and 2010

#Rename coloumn for dropout_age
names(dropout_age)[names(dropout_age) == "STATEFP"] <- "bpl"

# Load US state boundary data
us_states <- map_data("state")

#Covert to spatial data frama
maps_d_sf <- st_as_sf(maps_d_clean, coords = c("x_c", "y_c"))

# Filter dropout_age for 1980 and 2010
dropout_age_1980 <- filter(dropout_age, year == 1980)
dropout_age_2010 <- filter(dropout_age, year == 2010)

# Join the datasets mapd_d_sf and dropout_age_1980, as well as dropout_age_2010
maps_d_sf_1980 <- left_join(maps_d_sf, dropout_age_1980, by = c("bpl"))
maps_d_sf_2010 <- left_join(maps_d_sf, dropout_age_2010, by = c("bpl"))

#Convert the coloumn values to contain all lowercase values to match with 
# a us_states column
maps_d_sf_1980$NAME <- tolower(maps_d_sf_1980$NAME)
us_states_1980 <- left_join(us_states, maps_d_sf_1980, 
                            by =  c("region" = "NAME"))

maps_d_sf_2010$NAME <- tolower(maps_d_sf_2010$NAME)
us_states_2010 <- left_join(us_states, maps_d_sf_2010, 
                            by =  c("region" = "NAME"))

# Plots the map boundary
base_map <- ggplot() +
  geom_polygon(data = us_states_1980, aes(x = long, y = lat, group = group), 
               fill ="white", color = "grey") +
  coord_map() +
  theme_minimal() +
  labs(x = "Longitude", y = "Latitude") +
  ggtitle("Map of US States")

#Set the colours for the different ages
colours <- c("14" = "aquamarine2", "15" = "seagreen", "16" = "aquamarine3", 
             "17" = "aquamarine4", "18" = "chartreuse4", "NA" = "white")

# Plot dropout age data within state boundaries
data_map_1980 <- ggplot() +
  geom_sf(data = maps_d_sf_1980, aes(fill = as.factor(dropage)), size = 2) +
  geom_polygon(data = us_states_1980, aes(x = long, y = lat, group = group, 
                                          fill = as.factor(dropage)), 
               color = "grey") +
  theme_void() +
  scale_fill_manual(values = colours, 
                    guide = "legend", name = "Dropout Age  ",
                    na.value = "white") +
  labs(title = "Dropout Age by State (1980)") +
  theme(plot.title = element_text(size = 10, hjust = 0.4),
        legend.title = element_text(size = 10)) +
  coord_sf(xlim = c(-130, -60), ylim = c(25, 50))

# Plot dropout age data within state boundaries
data_map_2010 <- ggplot() +
  geom_sf(data = maps_d_sf_2010, aes(fill = as.factor(dropage)), size = 2) +
  geom_polygon(data = us_states_2010, aes(x = long, y = lat, group = group, 
                                          fill = as.factor(dropage)), 
               color = "grey") +
  theme_void() +
  scale_fill_manual(values = colours, 
                    guide = "legend", name = "Dropout Age  ",
                    na.value = "white") +
  labs(title = "Dropout Age by State (2010)") +
  theme(plot.title = element_text(size = 10, hjust = 0.4),
        legend.title = element_text(size = 10)) +
  coord_sf(xlim = c(-130, -60), ylim = c(25, 50))

# Show the combined plot
figure2 <- grid.arrange(data_map_1980, data_map_2010, ncol = 1)

In terms of CSL reforms, @fig-dropout illustrates the age individuals are required to be in school until they can dropout. This dropout age is shown for each state, starting at 14 until 18. Initially, in 1980, there was significant variability among states for the dropout age. 16% of states set the age under 16, 54% of the states set the age at 16 and 26% at ages 17-18. However, by 2010, we saw a notable shift in the mandatory dropout ages. Only one state, Montana, had the dropout age under 16 and 62% of states set the age to be 18. As you can see, there was a significant increase in educational interventions, in the form of CSL reforms, between the years 1980 and 2010. This reflects a broader societal emphasis on the importance of education and its impact on various social outcomes such as crime rates.

# Discussion {#sec-discussion}
